---
title: "SST Validation Project"
author: "Katie Nigro"
date: "2026-08-04"
output: html_document
editor_options: 
  chunk_output_type: console
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

#Load libraries
```{r}
library(ClimateNAr)
library(terra)
library(tidyverse)
library(ggplot2)
library(tidyterra)
library(pals)
```


#Web API

```{r, eval=FALSE}
#install.packages("httr2")
library(httr)
library(jsonlite)


############## for loop using GET with ClimateNA API, historic 1970's normal ######################
seedlots = pien_pts %>% 
  rename(el=elev, lon=long, ID1 = no, ID2 = Provenance) %>% 
  select(-source) %>% 
  filter(!ID2=="Plantation")
  

seedlot_clim_compiled = data.frame()
for(i in 1:nrow(seedlots)){
  normal_61_90 = GET("https://api.climatena.ca/api/cnaApi6/LatLonEl", query = list(ID1 = seedlots$ID1[i], ID2 = seedlots$ID2[i], lat=seedlots$lat[i], lon=seedlots$lon[i], el=seedlots$el[i], prd="Normal_1961_1990.nrm", varYSM="Y"))
  
  data=fromJSON(rawToChar(normal_61_90$content))
  
  seedlot_clim_compiled <- bind_rows(seedlot_clim_compiled, data)

}

seedlot_clim_compiled

###############################################################################
points_2020s = pien_pts %>% 
  rename(el=elev, lon=long, ID1 = no, ID2 = Provenance) %>% 
  select(-source) %>% 
  mutate(prd="8GCMs_ensemble_ssp245_2011-2040.gcm", varYSM="Y")
  

res_compiled_2020s = data.frame()
for(i in 1:nrow(points_2020s)){
  res = GET("https://api.climatena.ca/api/cnaApi6/LatLonEl", query = list(ID1 = points_2020s$ID1[i], ID2 = points_2020s$ID2[i], lat=points_2020s$lat[i], lon=points_2020s$lon[i], el=points_2020s$el[i], prd=points_2020s$prd[i], varYSM=points_2020s$varYSM[i]))
  
  data=fromJSON(rawToChar(res$content))
  
  res_compiled_2020s <- bind_rows(res_compiled_2020s, data)

}

res_compiled_2020s

```


Trying R package

```{r, eval=FALSE}
#install.packages('C:/Users/KatherineNigro/Box/01. katherine.nigro Workspace/ClimateNAr.zip', repos=NULL)


###crop DEM

dem <- "C:/Users/KatherineNigro/Box/01. katherine.nigro Workspace/spatial/GMTED2010N30W120_150/30n120w_20101117_gmted_mea150.tif"
us <- geodata::gadm(country = "USA",  level = 1, resolution = 2,
             path = "C:/Users/KatherineNigro/Box/01. katherine.nigro Workspace/PIENproject")
us_CO <- us[match("Colorado",us$NAME_1),]
  


dem_rast <- rast("C:/Users/KatherineNigro/Box/01. katherine.nigro Workspace/spatial/GMTED2010N30W120_150/30n120w_20101117_gmted_mea150.tif")
dem_crop_CO <- crop(dem_rast, ext(us_CO))
plot(dem_crop_CO)
crs(dem_crop_CO)

writeRaster(dem_crop_CO, "C:/Users/KatherineNigro/Box/01. katherine.nigro Workspace/spatial/30n120w_20101117_gmted_mea150_cropCO.tif")


###cropped
demToCSV("C:/Users/KatherineNigro/Box/01. katherine.nigro Workspace/spatial/GMTED2010N30W120_150/30n120w_20101117_gmted_mea150_cropCO.tif")

dem_csv_crop_CO <- read.csv("C:/Users/KatherineNigro/Box/01. katherine.nigro Workspace/spatial/GMTED2010N30W120_150/30n120w_20101117_gmted_mea150_cropCO.csv")
head(dem_csv_crop_CO)


###the whole thing
demToCSV(dem)

dem_csv <- read.csv("C:/Users/KatherineNigro/Box/01. katherine.nigro Workspace/spatial/GMTED2010N30W120_150/30n120w_20101117_gmted_mea150.csv")
head(dem_csv)


#using a CSV input file - cropped to CO
inputFile = "C:/Users/KatherineNigro/Box/01. katherine.nigro Workspace/spatial/30n120w_20101117_gmted_mea150_cropCO.tif"
varList=c('MCMT') # or varList='YS'
periodList=
c('Normal_1961_1990.nrm')
outDir= 'C:/Users/KatherineNigro/Box/01. katherine.nigro Workspace/spatial/'
test <- ClimateNAr::ClimateNAr(inputFile,periodList,varList,outDir) ##this took ~ 4 minutes for state of CO, 1 climate var, 1 normal period

pien_pts_normal_mcmt <- terra::extract(test, adjusted_provs, method="simple")

bind_cols(data.frame(adjusted_provs), data.frame(crds(adjusted_provs)),pien_pts_normal_mcmt)



```

# DEM
```{r}
#get DEM from GMTED 

dem_rast1 <- rast("C:/Users/KatherineNigro/Box/01. katherine.nigro Workspace/spatial/GMTED2010N30W120_150/30n120w_20101117_gmted_mea150.tif")
dem_rast2 <- rast("C:/Users/KatherineNigro/Box/01. katherine.nigro Workspace/spatial/GMTED2010N50W120_150/50n120w_20101117_gmted_mea150.tif")
dem_rast3 <- rast("C:/Users/KatherineNigro/Box/01. katherine.nigro Workspace/spatial/GMTED2010N30W150_150/30n150w_20101117_gmted_mea150.tif")
dem_rast4 <- rast("C:/Users/KatherineNigro/Box/01. katherine.nigro Workspace/spatial/GMTED2010N50W150_150/50n150w_20101117_gmted_mea150.tif")

dem_west <- mosaic(dem_rast1, dem_rast2, dem_rast3, dem_rast4, fun="first")

us <- geodata::gadm(country = "USA",  level = 1, resolution = 2,
             path = "C:/Users/KatherineNigro/Box/01. katherine.nigro Workspace/PIENproject")


ggplot()+
  geom_spatraster(data=dem_west)+
  geom_spatvector(data=us, fill=NA, color="black")
```

#Climate Match Fxn

```{r}
###make climate match function
#make sure all time periods needed are included here

climate_match <- function(prov_clim, plant_clim, mcmt_TL, shm_TL, plant_period){

m=prov_clim %>% select(provenance_name)

if(1140 %in% plant_period == TRUE){
mcmt_MS_1140 <- (abs(prov_clim$Normal_1961_1990_MCMT - plant_clim$`8GCMs_ensemble_ssp245_2011-2040_MCMT`)/mcmt_TL)^2
shm_MS_1140 <- (abs(prov_clim$Normal_1961_1990_SHM - plant_clim$`8GCMs_ensemble_ssp245_2011-2040_SHM`)/shm_TL)^2

m_1140 <- data.frame(m_1140 = (-1*(sqrt(mcmt_MS_1140+shm_MS_1140)-1))*100) %>% 
  bind_cols(prov_clim) %>% 
  bind_cols(delta_mcmt_1140 =  plant_clim$`8GCMs_ensemble_ssp245_2011-2040_MCMT` - prov_clim$Normal_1961_1990_MCMT) %>% 
  bind_cols(delta_shm_1140 = plant_clim$`8GCMs_ensemble_ssp245_2011-2040_SHM` - prov_clim$Normal_1961_1990_SHM)

m = left_join(m, m_1140)
}

if(9120 %in% plant_period == TRUE){
mcmt_MS_9120 <- (abs(prov_clim$Normal_1961_1990_MCMT - plant_clim$Normal_1991_2020_MCMT)/mcmt_TL)^2
shm_MS_9120 <- (abs(prov_clim$Normal_1961_1990_SHM - plant_clim$Normal_1991_2020_SHM)/shm_TL)^2

m_9120 <- data.frame(m_9120 = (-1*(sqrt(mcmt_MS_9120+shm_MS_9120)-1))*100) %>% 
  bind_cols(prov_clim) %>% 
  bind_cols(delta_mcmt_9120 =  plant_clim$Normal_1991_2020_MCMT - prov_clim$Normal_1961_1990_MCMT) %>% 
  bind_cols(delta_shm_9120 = plant_clim$Normal_1991_2020_SHM - prov_clim$Normal_1961_1990_SHM)

m = left_join(m, m_9120)
}

if(8110 %in% plant_period == TRUE){
mcmt_MS_8110 <- (abs(prov_clim$Normal_1961_1990_MCMT - plant_clim$Normal_1981_2010_MCMT)/mcmt_TL)^2
shm_MS_8110 <- (abs(prov_clim$Normal_1961_1990_SHM - plant_clim$Normal_1981_2010_SHM)/shm_TL)^2

m_8110 <- data.frame( m_8110 = (-1*(sqrt(mcmt_MS_8110+shm_MS_8110)-1))*100) %>% 
  bind_cols(prov_clim) %>% 
  bind_cols(delta_mcmt_8110 =  plant_clim$Normal_1981_2010_MCMT - prov_clim$Normal_1961_1990_MCMT) %>% 
  bind_cols(delta_shm_8110 = plant_clim$Normal_1981_2010_SHM - prov_clim$Normal_1961_1990_SHM)

m=left_join(m, m_8110)
}


if(7100 %in% plant_period == TRUE){
mcmt_MS_7100 <- (abs(prov_clim$Normal_1961_1990_MCMT - plant_clim$Normal_1971_2000_MCMT)/mcmt_TL)^2
shm_MS_7100 <- (abs(prov_clim$Normal_1961_1990_SHM - plant_clim$Normal_1971_2000_SHM)/shm_TL)^2

m_7100 <- data.frame( m_7100 = (-1*(sqrt(mcmt_MS_7100+shm_MS_7100)-1))*100) %>% 
  bind_cols(prov_clim) %>% 
  bind_cols(delta_mcmt_7100 =  plant_clim$Normal_1971_2000_MCMT - prov_clim$Normal_1961_1990_MCMT) %>% 
  bind_cols(delta_shm_7100 = plant_clim$Normal_1971_2000_SHM - prov_clim$Normal_1961_1990_SHM)

m=left_join(m, m_7100)
}


if(6190 %in% plant_period == TRUE){
mcmt_MS_6190 <- (abs(prov_clim$Normal_1961_1990_MCMT - plant_clim$Normal_1961_1990_MCMT)/mcmt_TL)^2
shm_MS_6190 <- (abs(prov_clim$Normal_1961_1990_SHM - plant_clim$Normal_1961_1990_SHM)/shm_TL)^2

m_6190 <- data.frame( m_6190 = (-1*(sqrt(mcmt_MS_6190+shm_MS_6190)-1))*100) %>% 
  bind_cols(prov_clim) %>% 
  bind_cols(delta_mcmt_6190 =  plant_clim$Normal_1961_1990_MCMT - prov_clim$Normal_1961_1990_MCMT) %>% 
  bind_cols(delta_shm_6190 = plant_clim$Normal_1961_1990_SHM - prov_clim$Normal_1961_1990_SHM)

m=left_join(m, m_6190)
}

return(m)}
```

#Run SST for lodgepole pine data

To run SST, we need 
1) Normal 1961 - 1990 ClimateNA data for each seedlot
2) Desired future projection time period and emissions scenario for the planting site 
    --> for Kat's study, let's say this is 1991 - 2020

##Study points

```{r}
pico_data <- read.csv("C:/Users/KatherineNigro/Box/01. katherine.nigro Workspace/SST_validation/data/PICO/warnick/Lodgepole_data_KW.csv") %>% 
  mutate(provenance_name=paste0(substr(provenance_name,0,2), provenance_elev_m))

pico_all_vect <- bind_rows(
  pico_data %>% 
  select(c(planting_site_name, planting_site_lat, planting_site_lon)) %>% 
  unique() %>% 
  rename(lat = planting_site_lat, lon = planting_site_lon, name=planting_site_name) %>% 
  mutate(site_type = "planting site")
,
pico_data %>% 
  select(c(provenance_name, provenance_lat, provenance_lon)) %>% 
  unique() %>% 
  rename(lat = provenance_lat, lon = provenance_lon, name=provenance_name) %>% 
  mutate(site_type="provenance")) %>% 
  filter(!is.na(lat)) %>% 
  vect(geom=c("lon","lat"), crs=crs(us))

writeVector(pico_all_vect, "C:/Users/KatherineNigro/Box/01. katherine.nigro Workspace/SST_validation/data/PICO/warnick/pico_all_vect.shp", overwrite=T)

pico_plantation_vect <- pico_data %>% 
  select(c(planting_site_name, planting_site_lat, planting_site_lon)) %>% 
  unique() %>% 
  filter(!is.na(planting_site_lat)) %>% 
  vect(geom=c("planting_site_lon","planting_site_lat"), crs=crs(us))

pico_provs_vect <- pico_data %>% 
  select(c(provenance_name, provenance_lat, provenance_lon)) %>% 
  unique() %>% 
  filter(!is.na(provenance_lat)) %>% 
  vect(geom=c("provenance_lon","provenance_lat"), crs=crs(us))

```

##buffer points and crop dem

```{r}
pico_all_buffered <- buffer(pico_all_vect, width=4000)
plot(pico_all_buffered)

dem_pico_pts_crop <- crop(dem_west, pico_all_buffered)
dem_pico_pts_mask <- mask(dem_pico_pts_crop, pico_all_buffered)

plot(dem_pico_pts_mask)

writeRaster(dem_pico_pts_mask, "C:/Users/KatherineNigro/Box/01. katherine.nigro Workspace/SST_validation/data/PICO/warnick/gmted_mea150_picocrop_bufferedpts.tif", overwrite=TRUE)

##get climatena data
inputFile = "C:/Users/KatherineNigro/Box/01. katherine.nigro Workspace/SST_validation/data/PICO/warnick/gmted_mea150_picocrop_bufferedpts.tif"
varList=c('MCMT','SHM') # or varList='YS'
periodList=
c('Normal_1961_1990.nrm','Normal_1991_2020.nrm')
outDir= 'C:/Users/KatherineNigro/Box/01. katherine.nigro Workspace/SST_validation/data/PICO/warnick'
pico_climatena <- ClimateNAr::ClimateNAr(inputFile,periodList,varList,outDir) ##YES! This is the way; <30sec.

```

##extract climate from points

```{r}
##get all TIF files

pico_tif_files <- list.files(
  path = "C:/Users/KatherineNigro/Box/01. katherine.nigro Workspace/SST_validation/data/PICO/warnick/gmted_mea150_picocrop_bufferedpts", 
  pattern = "\\.tif$", 
  recursive = TRUE, 
  full.names = TRUE
)

pico_climate_stack<- rast(pico_tif_files)
names(pico_climate_stack)<- paste(str_match(pico_tif_files, ".*/(.*)/.*$")[, 2],str_match(pico_tif_files, "/([^/]+)\\.tif$")[,2],sep="_")


pico_plantation_clim <- terra::extract(pico_climate_stack, pico_plantation_vect, method="simple") %>% 
  bind_cols(as.data.frame(pico_plantation_vect)) %>% 
  bind_cols(crds(pico_plantation_vect))
pico_provenance_clim <- terra::extract(pico_climate_stack, pico_provs_vect, method="simple") %>% 
  bind_cols(as.data.frame(pico_provs_vect)) %>% 
  bind_cols(crds(pico_provs_vect)) 

```

##calculate match scores

```{r}
##get match scores

pico_mcmt_shm_scores<- climate_match(pico_provenance_clim,pico_plantation_clim,mcmt_TL=2, shm_TL=pico_plantation_clim$Normal_1961_1990_SHM/2, plant_period=c(6190,9120)) %>% 
  select(provenance_name, m_6190, m_9120, everything())

pico_mcmt_shm_scores_long <- pico_mcmt_shm_scores %>% 
  pivot_longer(m_6190:m_9120, values_to = "m", names_to = "period") %>% 
  mutate(years=substr(period, 3,6))

pico_provs_by_lat <- pico_data %>% 
  select(provenance_name, provenance_lat) %>% 
  filter(!is.na(provenance_lat)) %>% 
  unique() %>% 
  arrange(provenance_lat)

ggplot(pico_mcmt_shm_scores_long, aes(x=as.numeric(years), y=m, col=factor(provenance_name, levels=pico_provs_by_lat$provenance_name)))+
  geom_point()+
  geom_line()+
  geom_hline(yintercept=0)
```

```{r, eval=FALSE}
###test in the online seedlot selection tool

pico_provs_forsst <- pico_data %>% 
  select(c(provenance_name, provenance_lat, provenance_lon)) %>% 
  unique() %>% 
  rename(Name = provenance_name, y=provenance_lat, x=provenance_lon)

write.csv(pico_provs_forsst, "C:/Users/KatherineNigro/Box/01. katherine.nigro Workspace/SST_validation/data/PICO/warnick/pico_provs_forsst.csv")

write.csv(pico_mcmt_shm_scores, "C:/Users/KatherineNigro/Box/01. katherine.nigro Workspace/SST_validation/data/PICO/warnick/pico_calculated_scores_mcmt_shm.csv")
```

##visualizing data

```{r}
pico_ht_dbh<- pico_data %>% 
  select(c(treeID, provenance_name, block, ht_cm, dbh_cm)) %>% 
  unique() %>% 
  left_join(pico_mcmt_shm_scores_long, by="provenance_name")

ggplot(pico_ht_dbh)+
  geom_boxplot(aes(x=provenance_name, y=ht_cm))

ggplot(pico_ht_dbh)+
  geom_point(aes(x=m, y=ht_cm, color=provenance_name), size=3)+
  xlab("match score")+
  ylab("height (m)")

ggplot(pico_ht_dbh %>% filter(m>0), aes(x=m, y=ht_cm))+
  geom_point(size=3)+
  geom_smooth(method="lm")+
  xlab("match score")+
  ylab("height (m)")

ggplot(pico_ht_dbh)+
  geom_point(aes(x=m, y=dbh_cm, color=provenance_name), size=3)+
  xlab("match score")+
  ylab("dbh (cm)")

ggplot(pico_ht_dbh %>% filter(m>0), aes(x=m, y=dbh_cm))+
  geom_point(size=3)+
  geom_smooth(method="lm")+
  xlab("match score")+
  ylab("height (m)")

```

#Run SST for Engelmann spruce data

To run SST, we need 
1) Normal 1961 - 1990 ClimateNA data for each seedlot
2) Desired future projection time period and emissions scenario for the planting site 
    --> for PIEN, we have height data for 1974, 1979, 1985, 1996, 2006, 2014 and DBH for 1991, 1996, 2006 and 2014
    --> so we will want the 3 climate normal periods 1971-2000, 1981-2010 & 1991-2020

##Study points

```{r}
pien_data <- read.csv("C:/Users/KatherineNigro/Box/01. katherine.nigro Workspace/SST_validation/data/PIEN/nigro/all_data_pien_forSST.csv")

pien_all_vect <- bind_rows(
  pien_data %>% 
  select(c(planting_site_name, planting_site_lat, planting_site_lon)) %>% 
  unique() %>% 
  rename(lat = planting_site_lat, lon = planting_site_lon, name=planting_site_name) %>% 
  mutate(site_type = "planting site")
,
pien_data %>% 
  select(c(provenance_name, provenance_lat, provenance_lon)) %>% 
  unique() %>% 
  rename(lat = provenance_lat, lon = provenance_lon, name=provenance_name) %>% 
  mutate(site_type="provenance")) %>% 
  filter(!is.na(lat)) %>% 
  vect(geom=c("lon","lat"), crs=crs(us))

#writeVector(pien_all_vect, "C:/Users/KatherineNigro/Box/01. katherine.nigro Workspace/SST_validation/data/PIEN/nigro/pien_all_vect.shp")

pien_plantation_vect <- pien_data %>% 
  select(c(planting_site_name, planting_site_lat, planting_site_lon)) %>% 
  unique() %>% 
  filter(!is.na(planting_site_lat)) %>% 
  vect(geom=c("planting_site_lon","planting_site_lat"), crs=crs(us))

pien_provs_vect <- pien_data %>% 
  select(c(provenance_name, provenance_lat, provenance_lon)) %>% 
  unique() %>% 
  filter(!is.na(provenance_lat)) %>% 
  vect(geom=c("provenance_lon","provenance_lat"), crs=crs(us))

```

##buffer points and crop dem

```{r}
pien_all_buffered <- buffer(pien_all_vect, width=4000)
plot(pien_all_buffered)

dem_pien_pts_crop <- crop(dem_west, pien_all_buffered)
dem_pien_pts_mask <- mask(dem_pien_pts_crop, pien_all_buffered)

plot(dem_pien_pts_mask)

writeRaster(dem_pien_pts_mask, "C:/Users/KatherineNigro/Box/01. katherine.nigro Workspace/SST_validation/data/PIEN/nigro/gmted_mea150_piencrop_bufferedpts.tif", overwrite=TRUE)

##get climatena data
inputFile = "C:/Users/KatherineNigro/Box/01. katherine.nigro Workspace/SST_validation/data/PIEN/nigro/gmted_mea150_piencrop_bufferedpts.tif"
varList=c('MCMT','SHM') # or varList='YS'
periodList=
c('Normal_1961_1990.nrm','Normal_1971_2000.nrm','Normal_1981_2010.nrm','Normal_1991_2020.nrm')
outDir= 'C:/Users/KatherineNigro/Box/01. katherine.nigro Workspace/SST_validation/data/PIEN/nigro'
pien_climatena <- ClimateNAr::ClimateNAr(inputFile,periodList,varList,outDir) ##YES! This is the way; <30sec.

```

##extract climate from points

```{r}
##get all TIF files

pien_tif_files <- list.files(
  path = "C:/Users/KatherineNigro/Box/01. katherine.nigro Workspace/SST_validation/data/PIEN/nigro/gmted_mea150_piencrop_bufferedpts", 
  pattern = "\\.tif$", 
  recursive = TRUE, 
  full.names = TRUE
)

pien_climate_stack<- rast(pien_tif_files)
names(pien_climate_stack)<- paste(str_match(pien_tif_files, ".*/(.*)/.*$")[, 2],str_match(pien_tif_files, "/([^/]+)\\.tif$")[,2],sep="_")


pien_plantation_clim <- terra::extract(pien_climate_stack, pien_plantation_vect, method="simple") %>% 
  bind_cols(as.data.frame(pien_plantation_vect)) %>% 
  bind_cols(crds(pien_plantation_vect))
pien_provenance_clim <- terra::extract(pien_climate_stack, pien_provs_vect, method="simple") %>% 
  bind_cols(as.data.frame(pien_provs_vect)) %>% 
  bind_cols(crds(pien_provs_vect)) 

```

##calculate match scores

```{r}
##get match scores

pien_mcmt_shm_scores<- climate_match(pien_provenance_clim,pien_plantation_clim,mcmt_TL=2, shm_TL=pien_plantation_clim$Normal_1961_1990_SHM/2, plant_period=c(6190,7100,8110,9120)) %>% 
  select(provenance_name, m_6190, m_7100, m_8110, m_9120, everything())

pien_mcmt_shm_scores_long <- pien_mcmt_shm_scores %>% 
  pivot_longer(m_6190:m_9120, values_to = "m", names_to = "period") %>% 
  mutate(years=substr(period, 3,6))

provs_by_lat <- pien_data %>% 
  select(provenance_name, provenance_lat) %>% 
  filter(!is.na(provenance_lat)) %>% 
  unique() %>% 
  arrange(provenance_lat)

ggplot(pien_mcmt_shm_scores_long, aes(x=as.numeric(years), y=m, col=factor(provenance_name, levels=provs_by_lat$provenance_name)))+
  geom_point()+
  geom_line()+
  scale_color_manual(values=unname(stepped()), name="Provenance")+
  geom_hline(yintercept=0)
```

```{r, eval=FALSE}
###test in the online seedlot selection tool

pien_provs_forsst <- pien_data %>% 
  select(c(provenance_name, provenance_lat, provenance_lon)) %>% 
  unique() %>% 
  rename(Name = provenance_name, y=provenance_lat, x=provenance_lon)

write.csv(pien_provs_forsst, "C:/Users/KatherineNigro/Box/01. katherine.nigro Workspace/SST_validation/data/PIEN/nigro/pien_provs_forsst.csv")

write.csv(match_scores_mcmt_shm, "C:/Users/KatherineNigro/Box/01. katherine.nigro Workspace/SST_validation/data/PIEN/nigro/pien_calculated_scores_mcmt_shm.csv")
```

##visualizing data

```{r}
pien_ht_dbh_2014<- pien_data %>% 
  filter(is.na(raw_ring_width_mm) & years_since_planting==44) %>% 
  select(c(treeID, provenance_name, block, ht_cm, dbh_cm)) %>% 
  unique() %>% 
  left_join(pien_mcmt_shm_scores, by="provenance_name")

ggplot(pien_ht_dbh_2014)+
  geom_boxplot(aes(x=provenance_name, y=ht_cm))

ggplot(pien_ht_dbh_2014)+
  geom_point(aes(x=m_9120, y=ht_cm, color=provenance_name), size=3)+
  xlab("match score")+
  ylab("height (m)")

ggplot(pien_ht_dbh_2014, aes(x=m_9120, y=ht_cm))+
  geom_point(size=3)+
  geom_smooth(method="lm")+
  xlab("match score")+
  ylab("height (m)")

ggplot(pien_ht_dbh_2014)+
  geom_point(aes(x=m_9120, y=dbh_cm, color=provenance_name), size=3)+
  xlab("match score")+
  ylab("dbh (cm)")

ggplot(pien_ht_dbh_2014, aes(x=m_9120, y=dbh_cm))+
  geom_point(size=3)+
  geom_smooth(method="lm")+
  xlab("match score")+
  ylab("dbh (cm)")

```


#Run SST for ponderosa pine Dixit data

To run SST, we need 
1) Normal 1961 - 1990 ClimateNA data for each seedlot
2) Desired future projection time period and emissions scenario for the planting site 
    --> for pipo, we have data for 2018-2025
    --> so we will want the climate normal period 2011-2040 under SSP245 (this is what ClimateNA recommends to use for the current period)

##Study points

```{r}
pipo_dixit_data <- read.csv("C:/Users/katherinenigro/Box/01. katherine.nigro Workspace/SST_validation/data/PIPO/dixit/Dixit_pipo_dixit_survival.csv")

head(pipo_dixit_data)

pipo_dixit_all_vect <- bind_rows(
  pipo_dixit_data %>% 
  select(c(planting_site_name, planting_site_lat, planting_site_lon)) %>% 
  unique() %>% 
  rename(lat = planting_site_lat, lon = planting_site_lon, name=planting_site_name) %>% 
  mutate(site_type = "planting site")
,
pipo_dixit_data %>% 
  select(c(provenance_name, provenance_lat, provenance_lon)) %>% 
  unique() %>% 
  rename(lat = provenance_lat, lon = provenance_lon, name=provenance_name) %>% 
  mutate(site_type="provenance")) %>% 
  filter(!is.na(lat)) %>% 
  vect(geom=c("lon","lat"), crs=crs(us))

#writeVector(pipo_dixit_all_vect, "C:/Users/katherinenigro/Box/01. katherine.nigro Workspace/SST_validation/data/PIPO/dixit/pipo_dixit_all_vect.shp")

pipo_dixit_plantation_vect <- pipo_dixit_data %>% 
  select(c(planting_site_name, planting_site_lat, planting_site_lon)) %>% 
  unique() %>% 
  filter(!is.na(planting_site_lat)) %>% 
  vect(geom=c("planting_site_lon","planting_site_lat"), crs=crs(us))

pipo_dixit_provs_vect <- pipo_dixit_data %>% 
  select(c(provenance_name, provenance_lat, provenance_lon)) %>% 
  unique() %>% 
  filter(!is.na(provenance_lat)) %>% 
  vect(geom=c("provenance_lon","provenance_lat"), crs=crs(us))

```

##buffer points and crop dem

```{r}
pipo_dixit_all_buffered <- buffer(pipo_dixit_all_vect, width=4000)
plot(pipo_dixit_all_buffered)

dem_pipo_dixit_pts_crop <- crop(dem_west, pipo_dixit_all_buffered)
dem_pipo_dixit_pts_mask <- mask(dem_pipo_dixit_pts_crop, pipo_dixit_all_buffered)

plot(dem_pipo_dixit_pts_mask)

writeRaster(dem_pipo_dixit_pts_mask, "C:/Users/katherinenigro/Box/01. katherine.nigro Workspace/SST_validation/data/PIPO/dixit/gmted_mea150_pipodixitcrop_bufferedpts.tif", overwrite=TRUE)

##get climatena data
inputFile = "C:/Users/katherinenigro/Box/01. katherine.nigro Workspace/SST_validation/data/PIPO/dixit/gmted_mea150_pipodixitcrop_bufferedpts.tif"
varList=c('MCMT','SHM') # or varList='YS'
periodList=
c('Normal_1961_1990.nrm','8GCMs_ensemble_ssp245_2011-2040.gcm')
outDir= 'C:/Users/katherinenigro/Box/01. katherine.nigro Workspace/SST_validation/data/PIPO/dixit'
pipo_dixit_climatena <- ClimateNAr::ClimateNAr(inputFile,periodList,varList,outDir) ##YES! This is the way; <30sec.

```

##extract climate from points

```{r}
##get all TIF files

pipo_dixit_tif_files <- list.files(
  path = "C:/Users/katherinenigro/Box/01. katherine.nigro Workspace/SST_validation/data/PIPO/dixit/gmted_mea150_pipodixitcrop_bufferedpts", 
  pattern = "\\.tif$", 
  recursive = TRUE, 
  full.names = TRUE
)

pipo_dixit_climate_stack<- rast(pipo_dixit_tif_files)
names(pipo_dixit_climate_stack)<- paste(str_match(pipo_dixit_tif_files, ".*/(.*)/.*$")[, 2],str_match(pipo_dixit_tif_files, "/([^/]+)\\.tif$")[,2],sep="_")


pipo_dixit_plantation_clim <- terra::extract(pipo_dixit_climate_stack, pipo_dixit_plantation_vect, method="simple") %>% 
  bind_cols(as.data.frame(pipo_dixit_plantation_vect)) %>% 
  bind_cols(crds(pipo_dixit_plantation_vect))
pipo_dixit_provenance_clim <- terra::extract(pipo_dixit_climate_stack, pipo_dixit_provs_vect, method="simple") %>% 
  bind_cols(as.data.frame(pipo_dixit_provs_vect)) %>% 
  bind_cols(crds(pipo_dixit_provs_vect)) 

```

##calculate match scores

```{r}
##get match scores

pipo_dixit_mcmt_shm_scores<- climate_match(pipo_dixit_provenance_clim,pipo_dixit_plantation_clim,mcmt_TL=2, shm_TL=40.3, plant_period=c(6190,1140)) %>% 
  select(provenance_name, m_6190, m_1140, everything())

pipo_dixit_mcmt_shm_scores_long <- pipo_dixit_mcmt_shm_scores %>% 
  pivot_longer(m_6190:m_1140, values_to = "m", names_to = "period") %>% 
  mutate(years=substr(period, 3,6))

provs_by_lat <- pipo_dixit_data %>% 
  select(provenance_name, provenance_lat) %>% 
  filter(!is.na(provenance_lat)) %>% 
  unique() %>% 
  arrange(provenance_lat)

ggplot(pipo_dixit_mcmt_shm_scores_long, aes(x=factor(years, levels=c("6190","1140")), y=m, col=factor(provenance_name, levels=provs_by_lat$provenance_name)))+
  geom_point()+
  geom_line(aes(group=provenance_name))+
  scale_color_manual(values=unname(stepped()), name="Provenance")+
  geom_hline(yintercept=0)
```

```{r, eval=FALSE}
###test in the online seedlot selection tool

pipo_dixit_provs_forsst <- pipo_dixit_data %>% 
  select(c(provenance_name, provenance_lat, provenance_lon)) %>% 
  unique() %>% 
  rename(Name = provenance_name, y=provenance_lat, x=provenance_lon)

write.csv(pipo_dixit_provs_forsst, "C:/Users/katherinenigro/Box/01. katherine.nigro Workspace/SST_validation/data/PIPO/dixit/pipo_dixit_provs_forsst.csv")

write.csv(pipo_dixit_mcmt_shm_scores, "C:/Users/katherinenigro/Box/01. katherine.nigro Workspace/SST_validation/data/PIPO/dixit/pipo_dixit_calculated_scores_mcmt_shm.csv")
```

##visualizing data

```{r}
pipo_dixit_surv<- pipo_dixit_data %>% 
  left_join(pipo_dixit_mcmt_shm_scores, by="provenance_name")

pipo_dixit_surv_sum<- pipo_dixit_surv %>% 
  group_by(provenance_name, years_since_planting, m_6190, m_1140) %>% 
  summarise(n=n(), status_sum=sum(status)) %>% #all have 48 trees to start
  mutate(surv.prop.from.start = status_sum/n)
  
ggplot(pipo_dixit_surv_sum, aes(x=m_1140, y=surv.prop.from.start))+
  geom_point()+
  geom_smooth(method="lm")+
  facet_wrap(~years_since_planting)
```


#Run SST for ponderosa pine Nigro data

To run SST, we need 
1) Normal 1961 - 1990 ClimateNA data for each seedlot
2) Desired future projection time period and emissions scenario for the planting site 
    --> for pipo, we have data for 2021-2025
    --> so we will want the climate normal period 2011-2040 under SSP245 (this is what ClimateNA recommends to use for the current period)

##Study points

```{r}
pipo_nigro_data <- read.csv("C:/Users/katherinenigro/Box/01. katherine.nigro Workspace/SST_validation/data/PIPO/nigro/planted_COpondo_forSST.csv") %>% 
  filter(!planting_site_name=="ML") ##removing ML since it had very different canopy conditions

head(pipo_nigro_data)

pipo_nigro_all_vect <- bind_rows(
  pipo_nigro_data %>% 
  select(c(planting_site_name, planting_site_lat, planting_site_lon)) %>% 
  unique() %>% 
  rename(lat = planting_site_lat, lon = planting_site_lon, name=planting_site_name) %>% 
  mutate(site_type = "planting site")
,
pipo_nigro_data %>% 
  select(c(provenance_name, provenance_lat, provenance_lon)) %>% 
  unique() %>% 
  rename(lat = provenance_lat, lon = provenance_lon, name=provenance_name) %>% 
  mutate(site_type="provenance")) %>% 
  filter(!is.na(lat)) %>% 
  vect(geom=c("lon","lat"), crs=crs(us))

#writeVector(pipo_nigro_all_vect, "C:/Users/katherinenigro/Box/01. katherine.nigro Workspace/SST_validation/data/PIPO/nigro/pipo_nigro_all_vect.shp")

pipo_nigro_plantation_vect <- pipo_nigro_data %>% 
  select(c(planting_site_name, planting_site_lat, planting_site_lon)) %>% 
  unique() %>% 
  filter(!is.na(planting_site_lat)) %>% 
  vect(geom=c("planting_site_lon","planting_site_lat"), crs=crs(us))

pipo_nigro_provs_vect <- pipo_nigro_data %>% 
  select(c(provenance_name, provenance_lat, provenance_lon)) %>% 
  unique() %>% 
  filter(!is.na(provenance_lat)) %>% 
  vect(geom=c("provenance_lon","provenance_lat"), crs=crs(us))

```

##buffer points and crop dem

```{r}
pipo_nigro_all_buffered <- buffer(pipo_nigro_all_vect, width=4000)
plot(pipo_nigro_all_buffered)

dem_pipo_nigro_pts_crop <- crop(dem_west, pipo_nigro_all_buffered)
dem_pipo_nigro_pts_mask <- mask(dem_pipo_nigro_pts_crop, pipo_nigro_all_buffered)

plot(dem_pipo_nigro_pts_mask)

writeRaster(dem_pipo_nigro_pts_mask, "C:/Users/katherinenigro/Box/01. katherine.nigro Workspace/SST_validation/data/PIPO/nigro/gmted_mea150_piponigrocrop_bufferedpts.tif", overwrite=TRUE)

##get climatena data
inputFile = "C:/Users/katherinenigro/Box/01. katherine.nigro Workspace/SST_validation/data/PIPO/nigro/gmted_mea150_piponigrocrop_bufferedpts.tif"
varList=c('MCMT','SHM') # or varList='YS'
periodList=
c('Normal_1961_1990.nrm','8GCMs_ensemble_ssp245_2011-2040.gcm')
outDir= 'C:/Users/katherinenigro/Box/01. katherine.nigro Workspace/SST_validation/data/PIPO/nigro'
pipo_nigro_climatena <- ClimateNAr::ClimateNAr(inputFile,periodList,varList,outDir) ##YES! This is the way; <30sec.

```

##extract climate from points

```{r}
##get all TIF files

pipo_nigro_tif_files <- list.files(
  path = "C:/Users/katherinenigro/Box/01. katherine.nigro Workspace/SST_validation/data/PIPO/nigro/gmted_mea150_piponigrocrop_bufferedpts", 
  pattern = "\\.tif$", 
  recursive = TRUE, 
  full.names = TRUE
)

pipo_nigro_climate_stack<- rast(pipo_nigro_tif_files)
names(pipo_nigro_climate_stack)<- paste(str_match(pipo_nigro_tif_files, ".*/(.*)/.*$")[, 2],str_match(pipo_nigro_tif_files, "/([^/]+)\\.tif$")[,2],sep="_")

pipo_nigro_plantation_clim <- terra::extract(pipo_nigro_climate_stack, pipo_nigro_plantation_vect, method="simple") %>% 
  bind_cols(as.data.frame(pipo_nigro_plantation_vect)) %>% 
  bind_cols(crds(pipo_nigro_plantation_vect))
pipo_nigro_provenance_clim <- terra::extract(pipo_nigro_climate_stack, pipo_nigro_provs_vect, method="simple") %>% 
  bind_cols(as.data.frame(pipo_nigro_provs_vect)) %>% 
  bind_cols(crds(pipo_nigro_provs_vect)) 

```

##calculate match scores

```{r}
##get match scores

pipo_nigro_mcmt_shm_scores_HH<- climate_match(pipo_nigro_provenance_clim,pipo_nigro_plantation_clim %>% filter(planting_site_name=="HH"),mcmt_TL=2, shm_TL = pipo_nigro_plantation_clim %>% filter(planting_site_name=="HH") %>% pull(Normal_1961_1990_SHM)/2, plant_period=c(1140)) %>% 
  select(provenance_name, m_1140, everything())

pipo_nigro_mcmt_shm_scores_MH<- climate_match(pipo_nigro_provenance_clim,pipo_nigro_plantation_clim %>% filter(planting_site_name=="MH"),mcmt_TL=2, shm_TL=pipo_nigro_plantation_clim %>% filter(planting_site_name=="MH") %>% pull(Normal_1961_1990_SHM)/2, plant_period=c(1140)) %>% 
  select(provenance_name, m_1140, everything())

###HH
pipo_nigro_mcmt_shm_scores_HH_long <- pipo_nigro_mcmt_shm_scores_HH %>% 
  pivot_longer(m_1140, values_to = "m", names_to = "period") %>% 
  mutate(years=substr(period, 3,6))

###MH
pipo_nigro_mcmt_shm_scores_MH_long <- pipo_nigro_mcmt_shm_scores_MH %>% 
  pivot_longer(m_1140, values_to = "m", names_to = "period") %>% 
  mutate(years=substr(period, 3,6))

pipo_nigro_provs_by_lat <- pipo_nigro_data %>% 
  select(provenance_name, provenance_lat) %>% 
  filter(!is.na(provenance_lat)) %>% 
  unique() %>% 
  arrange(provenance_lat)

ggplot(pipo_nigro_mcmt_shm_scores_HH_long, aes(x=factor(provenance_name, levels=pipo_nigro_provs_by_lat$provenance_name), y=m, fill=factor(provenance_name, levels=pipo_nigro_provs_by_lat$provenance_name)))+
  geom_bar(stat="identity")+
  geom_hline(yintercept=0)

ggplot(pipo_nigro_mcmt_shm_scores_MH_long, aes(x=factor(provenance_name, levels=pipo_nigro_provs_by_lat$provenance_name), y=m, fill=factor(provenance_name, levels=pipo_nigro_provs_by_lat$provenance_name)))+
  geom_bar(stat="identity")+
  geom_hline(yintercept=0)

```

```{r, eval=FALSE}
###test in the online seedlot selection tool

pipo_nigro_provs_forsst <- pipo_nigro_data %>% 
  select(c(provenance_name, provenance_lat, provenance_lon)) %>% 
  unique() %>% 
  rename(Name = provenance_name, y=provenance_lat, x=provenance_lon)

write.csv(pipo_nigro_provs_forsst, "C:/Users/katherinenigro/Box/01. katherine.nigro Workspace/SST_validation/data/PIPO/nigro/pipo_nigro_provs_forsst.csv")

#write.csv(pipo_nigro_mcmt_shm_scores, "C:/Users/katherinenigro/Box/01. katherine.nigro Workspace/SST_validation/data/PIPO/nigro/pipo_nigro_calculated_scores_mcmt_shm.csv")
```

##visualizing data

###MH
```{r}
pipo_nigro_ht_gld_yr4<- pipo_nigro_data %>% 
  filter(years_since_planting==4 & planting_site_name=="MH") %>% 
  select(c(treeID, provenance_name, block, ht_cm, gld_cm)) %>% 
  unique() %>% 
  left_join(pipo_nigro_mcmt_shm_scores_MH, by="provenance_name")

ggplot(pipo_nigro_ht_gld_yr4)+
  geom_boxplot(aes(x=provenance_name, y=ht_cm))

ggplot(pipo_nigro_ht_gld_yr4)+
  geom_point(aes(x=m_1140, y=ht_cm, color=provenance_name), size=3)+
  xlab("match score")+
  ylab("height (m)")

ggplot(pipo_nigro_ht_gld_yr4, aes(x=m_1140, y=ht_cm))+
  geom_point(size=3)+
  geom_smooth(method="lm")+
  xlab("match score")+
  ylab("height (m)")

ggplot(pipo_nigro_ht_gld_yr4)+
  geom_point(aes(x=m_1140, y=gld_cm, color=provenance_name), size=3)+
  xlab("match score")+
  ylab("gld (cm)")

ggplot(pipo_nigro_ht_gld_yr4, aes(x=m_1140, y=gld_cm))+
  geom_point(size=3)+
  geom_smooth(method="lm")+
  xlab("match score")+
  ylab("gld (cm)")


pipo_nigro_data_init_4yr_ht<- pipo_nigro_data %>% filter(years_since_planting%in%c("initial","4") & planting_site_name=="MH") %>% 
  select(c(treeID, provenance_name, planting_site_name, years_since_planting, ht_cm)) %>% 
  pivot_wider(names_from = years_since_planting, values_from = ht_cm)

ggplot(pipo_nigro_data_init_4yr_ht, aes(x=initial, y=`4`))+
  geom_point()+
  geom_smooth(method="lm")

```


#Run SST for pinyon pine Redmond data

To run SST, we need 
1) Normal 1961 - 1990 ClimateNA data for each seedlot
2) Desired future projection time period and emissions scenario for the planting site 
    --> for pied, we have data from 2020s
    --> so we will want the climate normal period 2011-2040 under SSP245 (this is what ClimateNA recommends to use for the current period)

##Study points

```{r}
pied_redmond_data <- read.csv("C:/Users/katherinenigro/Box/01. katherine.nigro Workspace/SST_validation/data/PIED/redmond/PUR_germdat_NigroSST_EMH_Aug2026.csv")

head(pied_redmond_data)

pied_redmond_all_vect <- bind_rows(
  pied_redmond_data %>% 
  select(c(planting_site_name, planting_site_lat, planting_site_lon)) %>% 
  unique() %>% 
  rename(lat = planting_site_lat, lon = planting_site_lon, name=planting_site_name) %>% 
  mutate(site_type = "planting site")
,
pied_redmond_data %>% 
  select(c(provenance_name, provenance_lat, provenance_lon)) %>% 
  mutate(provenance_name=substr(provenance_name, 0,3)) %>% 
  unique() %>% 
  rename(lat = provenance_lat, lon = provenance_lon, name=provenance_name) %>% 
  mutate(site_type="provenance")) %>% 
  filter(!is.na(lat)) %>% 
  vect(geom=c("lon","lat"), crs=crs(us))

#writeVector(pied_redmond_all_vect, "C:/Users/katherinenigro/Box/01. katherine.nigro Workspace/SST_validation/data/PIED/redmond/pied_redmond_all_vect.shp", overwrite=TRUE)

pied_redmond_plantation_vect <- pied_redmond_data %>% 
  select(c(planting_site_name, planting_site_lat, planting_site_lon)) %>% 
  unique() %>% 
  filter(!is.na(planting_site_lat)) %>% 
  vect(geom=c("planting_site_lon","planting_site_lat"), crs=crs(us))

pied_redmond_provs_vect <- pied_redmond_data %>% 
  select(c(provenance_name, provenance_lat, provenance_lon)) %>% 
    mutate(provenance_name=substr(provenance_name, 0,3)) %>% 
  unique() %>% 
  filter(!is.na(provenance_lat)) %>% 
  vect(geom=c("provenance_lon","provenance_lat"), crs=crs(us))

```

##buffer points and crop dem

```{r}
pied_redmond_all_buffered <- buffer(pied_redmond_all_vect, width=4000)
plot(pied_redmond_all_buffered)

dem_pied_redmond_pts_crop <- crop(dem_west, pied_redmond_all_buffered)
dem_pied_redmond_pts_mask <- mask(dem_pied_redmond_pts_crop, pied_redmond_all_buffered)

plot(dem_pied_redmond_pts_mask)

writeRaster(dem_pied_redmond_pts_mask, "C:/Users/katherinenigro/Box/01. katherine.nigro Workspace/SST_validation/data/PIED/redmond/gmted_mea150_piedredmondcrop_bufferedpts.tif", overwrite=TRUE)

##get climatena data
inputFile = "C:/Users/katherinenigro/Box/01. katherine.nigro Workspace/SST_validation/data/PIED/redmond/gmted_mea150_piedredmondcrop_bufferedpts.tif"
varList=c('MCMT','SHM') # or varList='YS'
periodList=
c('Normal_1961_1990.nrm','8GCMs_ensemble_ssp245_2011-2040.gcm')
outDir= 'C:/Users/katherinenigro/Box/01. katherine.nigro Workspace/SST_validation/data/PIED/redmond'
pied_redmond_climatena <- ClimateNAr::ClimateNAr(inputFile,periodList,varList,outDir) ##YES! This is the way; <30sec.

```

##extract climate from points

```{r}
##get all TIF files

pied_redmond_tif_files <- list.files(
  path = "C:/Users/katherinenigro/Box/01. katherine.nigro Workspace/SST_validation/data/PIED/redmond/gmted_mea150_piedredmondcrop_bufferedpts", 
  pattern = "\\.tif$", 
  recursive = TRUE, 
  full.names = TRUE
)

pied_redmond_climate_stack<- rast(pied_redmond_tif_files)
names(pied_redmond_climate_stack)<- paste(str_match(pied_redmond_tif_files, ".*/(.*)/.*$")[, 2],str_match(pied_redmond_tif_files, "/([^/]+)\\.tif$")[,2],sep="_")

pied_redmond_plantation_clim <- terra::extract(pied_redmond_climate_stack, pied_redmond_plantation_vect, method="simple") %>% 
  bind_cols(as.data.frame(pied_redmond_plantation_vect)) %>% 
  bind_cols(crds(pied_redmond_plantation_vect))
pied_redmond_provenance_clim <- terra::extract(pied_redmond_climate_stack, pied_redmond_provs_vect, method="simple") %>% 
  bind_cols(as.data.frame(pied_redmond_provs_vect)) %>% 
  bind_cols(crds(pied_redmond_provs_vect)) 

```

##calculate match scores

```{r}
##get match scores for each planting site

#NW
pied_redmond_mcmt_shm_scores_NWL<- climate_match(pied_redmond_provenance_clim,pied_redmond_plantation_clim %>% filter(planting_site_name=="NWL"),mcmt_TL=2, shm_TL=pied_redmond_plantation_clim %>% filter(planting_site_name=="NWL") %>% pull(Normal_1961_1990_SHM)/2, plant_period=c(1140)) %>% 
  select(provenance_name, m_1140, everything())

pied_redmond_mcmt_shm_scores_NWM<- climate_match(pied_redmond_provenance_clim,pied_redmond_plantation_clim %>% filter(planting_site_name=="NWM"),mcmt_TL=2, shm_TL=pied_redmond_plantation_clim %>% filter(planting_site_name=="NWM") %>% pull(Normal_1961_1990_SHM)/2, plant_period=c(1140)) %>% 
  select(provenance_name, m_1140, everything())

#SW
pied_redmond_mcmt_shm_scores_SWL<- climate_match(pied_redmond_provenance_clim,pied_redmond_plantation_clim %>% filter(planting_site_name=="SWL"),mcmt_TL=2, shm_TL=pied_redmond_plantation_clim %>% filter(planting_site_name=="SWL") %>% pull(Normal_1961_1990_SHM)/2, plant_period=c(1140)) %>% 
  select(provenance_name, m_1140, everything())

pied_redmond_mcmt_shm_scores_SWM<- climate_match(pied_redmond_provenance_clim,pied_redmond_plantation_clim %>% filter(planting_site_name=="SWM"),mcmt_TL=2, shm_TL=pied_redmond_plantation_clim %>% filter(planting_site_name=="SWM") %>% pull(Normal_1961_1990_SHM)/2, plant_period=c(1140)) %>% 
  select(provenance_name, m_1140, everything())

pied_redmond_mcmt_shm_scores_SWH<- climate_match(pied_redmond_provenance_clim,pied_redmond_plantation_clim %>% filter(planting_site_name=="SWH"),mcmt_TL=2, shm_TL=pied_redmond_plantation_clim %>% filter(planting_site_name=="SWH") %>% pull(Normal_1961_1990_SHM)/2, plant_period=c(1140)) %>% 
  select(provenance_name, m_1140, everything())

#AZ
pied_redmond_mcmt_shm_scores_AZL<- climate_match(pied_redmond_provenance_clim,pied_redmond_plantation_clim %>% filter(planting_site_name=="AZL"),mcmt_TL=2, shm_TL=pied_redmond_plantation_clim %>% filter(planting_site_name=="AZL") %>% pull(Normal_1961_1990_SHM)/2, plant_period=c(1140)) %>% 
  select(provenance_name, m_1140, everything())

pied_redmond_mcmt_shm_scores_AZM<- climate_match(pied_redmond_provenance_clim,pied_redmond_plantation_clim %>% filter(planting_site_name=="AZM"),mcmt_TL=2, shm_TL=pied_redmond_plantation_clim %>% filter(planting_site_name=="AZM") %>% pull(Normal_1961_1990_SHM)/2, plant_period=c(1140)) %>% 
  select(provenance_name, m_1140, everything())
```


```{r}

pied_redmond_mcmt_shm_scores_long <- rbind(
#NW
pied_redmond_mcmt_shm_scores_NWL %>% mutate(planting_site_name="NWL"),
pied_redmond_mcmt_shm_scores_NWM %>% mutate(planting_site_name="NWM"),
#SW
pied_redmond_mcmt_shm_scores_SWL %>% mutate(planting_site_name="SWL"),
pied_redmond_mcmt_shm_scores_SWM %>% mutate(planting_site_name="SWM"),
pied_redmond_mcmt_shm_scores_SWH %>% mutate(planting_site_name="SWH"),
#AZ
pied_redmond_mcmt_shm_scores_AZL %>% mutate(planting_site_name="AZL"),
pied_redmond_mcmt_shm_scores_AZM %>% mutate(planting_site_name="AZM")
) %>% 
  pivot_longer(m_1140, names_to="years", values_to = "m")

pied_redmond_provs_by_lat <- pied_redmond_data %>% 
  select(provenance_name, provenance_lat) %>% 
  mutate(provenance_name=substr(provenance_name, 0,3)) %>% 
  filter(!is.na(provenance_lat)) %>% 
  unique() %>% 
  arrange(provenance_lat)

ggplot(pied_redmond_mcmt_shm_scores_long, aes(x=factor(provenance_name, levels=pied_redmond_provs_by_lat$provenance_name), y=m, fill=factor(provenance_name, levels=pied_redmond_provs_by_lat$provenance_name)))+
  geom_bar(stat="identity")+
  geom_hline(yintercept=0)+
  scale_fill_manual(values=seq(1,5,1), name="Provenance")+
  facet_wrap(~planting_site_name)+
  xlab("")

```

```{r, eval=FALSE}
###test in the online seedlot selection tool

pied_redmond_provs_forsst <- pied_redmond_data %>% 
  select(c(provenance_name, provenance_lat, provenance_lon)) %>% 
  unique() %>% 
  rename(Name = provenance_name, y=provenance_lat, x=provenance_lon)

write.csv(pied_redmond_provs_forsst, "C:/Users/katherinenigro/Box/01. katherine.nigro Workspace/SST_validation/data/PIED/redmond/pied_redmond_provs_forsst.csv")

#write.csv(pied_redmond_mcmt_shm_scores, "C:/Users/katherinenigro/Box/01. katherine.nigro Workspace/SST_validation/data/PIED/redmond/pied_redmond_calculated_scores_mcmt_shm.csv")
```

##visualizing data

```{r}
pied_redmond_surv <- pied_redmond_data %>% 
  mutate(provenance_name=substr(provenance_name, 0,3)) %>% 
  group_by(provenance_name, planting_site_name, years_since_planting) %>% 
    summarise(n_surv=sum(number_surviving), n_tot=n()) %>% 
  mutate(surv.prop = n_surv/n_tot) %>% 
  left_join(pied_redmond_mcmt_shm_scores_long)

ggplot(data=pied_redmond_surv, aes(x=provenance_name, y=surv.prop))+
  geom_bar(stat="identity")+
  facet_wrap(~planting_site_name)

ggplot(data=pied_redmond_surv %>% filter(years_since_planting==2), aes(x=m, y=surv.prop))+
  geom_point()+
  facet_wrap(~planting_site_name)


```


#Run SST for pinyon pine Urza data

To run SST, we need 
1) Normal 1961 - 1990 ClimateNA data for each seedlot
2) Desired future projection time period and emissions scenario for the planting site 
    --> for pimo, we have data from 2020s
    --> so we will want the climate normal period 2011-2040 under SSP245 (this is what ClimateNA recommends to use for the current period)

##Study points

```{r}
pimo_urza_data <- read.csv("C:/Users/katherinenigro/Box/01. katherine.nigro Workspace/SST_validation/data/PIMO/urza/Urza_P.monophylla_data_forSST_seedlingdata.csv")

head(pimo_urza_data)

pimo_urza_all_vect <- bind_rows(
  pimo_urza_data %>% 
  select(c(planting_site_name, planting_site_lat, planting_site_lon)) %>% 
  unique() %>% 
  rename(lat = planting_site_lat, lon = planting_site_lon, name=planting_site_name) %>% 
  mutate(site_type = "planting site")
,
pimo_urza_data %>% 
  select(c(provenance_name, provenance_lat, provenance_lon)) %>% 
  unique() %>% 
  rename(lat = provenance_lat, lon = provenance_lon, name=provenance_name) %>% 
  mutate(site_type="provenance")) %>% 
  filter(!is.na(lat)) %>% 
  vect(geom=c("lon","lat"), crs=crs(us))

#writeVector(pimo_urza_all_vect, "C:/Users/katherinenigro/Box/01. katherine.nigro Workspace/SST_validation/data/PIMO/urza/pimo_urza_all_vect.shp", overwrite=TRUE)

pimo_urza_plantation_vect <- pimo_urza_data %>% 
  select(c(planting_site_name, planting_site_lat, planting_site_lon)) %>% 
  unique() %>% 
  filter(!is.na(planting_site_lat)) %>% 
  vect(geom=c("planting_site_lon","planting_site_lat"), crs=crs(us))

pimo_urza_provs_vect <- pimo_urza_data %>% 
  select(c(provenance_name, provenance_lat, provenance_lon)) %>% 
  unique() %>% 
  filter(!is.na(provenance_lat)) %>% 
  vect(geom=c("provenance_lon","provenance_lat"), crs=crs(us))

```

##buffer points and crop dem

```{r}
pimo_urza_all_buffered <- buffer(pimo_urza_all_vect, width=4000)
plot(pimo_urza_all_buffered)

dem_pimo_urza_pts_crop <- crop(dem_west, pimo_urza_all_buffered)
dem_pimo_urza_pts_mask <- mask(dem_pimo_urza_pts_crop, pimo_urza_all_buffered)

plot(dem_pimo_urza_pts_mask)

writeRaster(dem_pimo_urza_pts_mask, "C:/Users/katherinenigro/Box/01. katherine.nigro Workspace/SST_validation/data/PIMO/urza/gmted_mea150_pimourzacrop_bufferedpts.tif", overwrite=TRUE)

##get climatena data
inputFile = "C:/Users/katherinenigro/Box/01. katherine.nigro Workspace/SST_validation/data/PIMO/urza/gmted_mea150_pimourzacrop_bufferedpts.tif"
varList=c('MCMT','SHM') # or varList='YS'
periodList=
c('Normal_1961_1990.nrm','8GCMs_ensemble_ssp245_2011-2040.gcm')
outDir= 'C:/Users/katherinenigro/Box/01. katherine.nigro Workspace/SST_validation/data/PIMO/urza'
pimo_urza_climatena <- ClimateNAr::ClimateNAr(inputFile,periodList,varList,outDir) ##YES! This is the way; <30sec.

```

##extract climate from points

```{r}
##get all TIF files

pimo_urza_tif_files <- list.files(
  path = "C:/Users/katherinenigro/Box/01. katherine.nigro Workspace/SST_validation/data/PIMO/urza/gmted_mea150_pimourzacrop_bufferedpts", 
  pattern = "\\.tif$", 
  recursive = TRUE, 
  full.names = TRUE
)

pimo_urza_climate_stack<- rast(pimo_urza_tif_files)
names(pimo_urza_climate_stack)<- paste(str_match(pimo_urza_tif_files, ".*/(.*)/.*$")[, 2],str_match(pimo_urza_tif_files, "/([^/]+)\\.tif$")[,2],sep="_")

pimo_urza_plantation_clim <- terra::extract(pimo_urza_climate_stack, pimo_urza_plantation_vect, method="simple") %>% 
  bind_cols(as.data.frame(pimo_urza_plantation_vect)) %>% 
  bind_cols(crds(pimo_urza_plantation_vect))
pimo_urza_provenance_clim <- terra::extract(pimo_urza_climate_stack, pimo_urza_provs_vect, method="simple") %>% 
  bind_cols(as.data.frame(pimo_urza_provs_vect)) %>% 
  bind_cols(crds(pimo_urza_provs_vect)) 

```

##calculate match scores

```{r}
##get match scores for each planting site

#PUD-H
pimo_urza_mcmt_shm_scores_PUDH<- climate_match(pimo_urza_provenance_clim,pimo_urza_plantation_clim %>% filter(planting_site_name=="PUD-H"),mcmt_TL=2, shm_TL=pimo_urza_plantation_clim %>% filter(planting_site_name=="PUD-H") %>% pull(Normal_1961_1990_SHM)/2, plant_period=c(1140)) %>% 
  select(provenance_name, m_1140, everything())

#PUD-L
pimo_urza_mcmt_shm_scores_PUDL<- climate_match(pimo_urza_provenance_clim,pimo_urza_plantation_clim %>% filter(planting_site_name=="PUD-L"),mcmt_TL=2, shm_TL=pimo_urza_plantation_clim %>% filter(planting_site_name=="PUD-L") %>% pull(Normal_1961_1990_SHM)/2, plant_period=c(1140)) %>% 
  select(provenance_name, m_1140, everything())

#PUD_PL
pimo_urza_mcmt_shm_scores_PUDPL<- climate_match(pimo_urza_provenance_clim,pimo_urza_plantation_clim %>% filter(planting_site_name=="PUD_PL"),mcmt_TL=2, shm_TL=pimo_urza_plantation_clim %>% filter(planting_site_name=="PUD_PL") %>% pull(Normal_1961_1990_SHM)/2, plant_period=c(1140)) %>% 
  select(provenance_name, m_1140, everything())

```


```{r}

pimo_urza_mcmt_shm_scores_long <- rbind(
pimo_urza_mcmt_shm_scores_PUDH %>% mutate(planting_site_name="PUD-H"),
pimo_urza_mcmt_shm_scores_PUDL %>% mutate(planting_site_name="PUD-L"),
pimo_urza_mcmt_shm_scores_PUDPL %>% mutate(planting_site_name="PUD_PL")
) %>% 
  pivot_longer(m_1140, names_to="years", values_to = "m")

pimo_urza_provs_by_lat <- pimo_urza_data %>% 
  select(provenance_name, provenance_lat) %>% 
  filter(!is.na(provenance_lat)) %>% 
  unique() %>% 
  arrange(provenance_lat)

ggplot(pimo_urza_mcmt_shm_scores_long, aes(x=factor(provenance_name, levels=pimo_urza_provs_by_lat$provenance_name), y=m, fill=factor(provenance_name, levels=pimo_urza_provs_by_lat$provenance_name)))+
  geom_bar(stat="identity")+
  geom_hline(yintercept=0)+
  scale_fill_manual(values=c("#a9b836","#7562cd","#62be4a","#c454ba","#4e8c37","#d34782","#62c185","#cf4148","#41c0c7","#d1592a","#608acd","#dc9c36","#b47aba","#8f882d","#b35a6d","#388661","#e48c70","#696e2e","#9d622e","#beb16b"), name="Provenance")+
  facet_wrap(~planting_site_name)+
  xlab("")

```

```{r, eval=FALSE}
###test in the online seedlot selection tool

pimo_urza_provs_forsst <- pimo_urza_data %>% 
  select(c(provenance_name, provenance_lat, provenance_lon)) %>% 
  unique() %>% 
  rename(Name = provenance_name, y=provenance_lat, x=provenance_lon)

write.csv(pimo_urza_provs_forsst, "C:/Users/katherinenigro/Box/01. katherine.nigro Workspace/SST_validation/data/PIMO/urza/pimo_urza_provs_forsst.csv")

#write.csv(pimo_urza_mcmt_shm_scores, "C:/Users/katherinenigro/Box/01. katherine.nigro Workspace/SST_validation/data/PIMO/urza/pimo_urza_calculated_scores_mcmt_shm.csv")
```

##visualizing data

```{r}
pimo_urza_surv <- pimo_urza_data %>% 
  filter(!is.na(status)) %>%  ##gets rid of seedlings that were not surveyed
  group_by(provenance_name, planting_site_name, years_since_planting) %>% 
    summarise(n_surv=sum(status), n_tot=n()) %>% 
  mutate(surv.prop = n_surv/n_tot) %>% 
  left_join(pimo_urza_mcmt_shm_scores_long)

ggplot(data=pimo_urza_surv, aes(x=provenance_name, y=surv.prop, fill=provenance_name))+
  geom_bar(stat="identity", position = "dodge")+
  facet_grid(years_since_planting~planting_site_name)

ggplot(data=pimo_urza_surv %>% filter(years_since_planting>4), aes(x=m, y=surv.prop))+
  geom_point()+
  geom_smooth(method="lm")+
  facet_wrap(~planting_site_name)

pimo_urza_comb<- pimo_urza_data %>% 
  left_join(pimo_urza_mcmt_shm_scores_long)

ggplot(pimo_urza_comb %>% filter(!is.na(height_cm)), aes(x=m, y=height_cm))+
  geom_point()+
  facet_grid(years_since_planting~planting_site_name)+
  geom_vline(xintercept=0, color="red")
```



#Try AdaptWest data download

```{r, eval=FALSE}

normal6190_mcmt <- rast("C:/Users/KatherineNigro/Box/01. katherine.nigro Workspace/SST_validation/Normal_1961_1990_bioclim/Normal_1961_1990_MCMT.tif")

normal6190_shm <- rast("C:/Users/KatherineNigro/Box/01. katherine.nigro Workspace/SST_validation/Normal_1961_1990_bioclim/Normal_1961_1990_SHM.tif")

normal9120_mcmt <- rast("C:/Users/KatherineNigro/Box/01. katherine.nigro Workspace/SST_validation/Normal_1991_2020_bioclim/Normal_1991_2020_MCMT.tif")

normal9120_shm <- rast("C:/Users/KatherineNigro/Box/01. katherine.nigro Workspace/SST_validation/Normal_1991_2020_bioclim/Normal_1991_2020_SHM.tif")

future1140_mcmt <- rast("C:/Users/KatherineNigro/Box/01. katherine.nigro Workspace/SST_validation/ensemble_8GCMs_ssp245_2011_2040_bioclim/ensemble_8GCMs_ssp245_2011_2040_MCMT.tif")

future1140_shm <- rast("C:/Users/KatherineNigro/Box/01. katherine.nigro Workspace/SST_validation/ensemble_8GCMs_ssp245_2011_2040_bioclim/ensemble_8GCMs_ssp245_2011_2040_SHM.tif")

AW_clim_stack<- c(normal6190_mcmt,normal6190_shm, normal9120_mcmt, normal9120_shm, future1140_mcmt, future1140_shm)
names(AW_clim_stack)<-  c("Normal_1961_1990_MCMT","Normal_1961_1990_SHM", "Normal_1991_2020_MCMT", "Normal_1991_2020_SHM", "8GCMs_ensemble_ssp245_2011-2040_MCMT", "8GCMs_ensemble_ssp245_2011-2040_SHM")
```

##pico
```{r}
AW_clim_pico <- terra::extract(AW_clim_stack, pico_all_vect, method="simple") %>% 
  bind_cols(as.data.frame(pico_all_vect)) %>% 
  bind_cols(crds(pico_all_vect)) 

AW_clim_pico_plantation<- AW_clim_pico %>% 
  filter(site_type=="planting site") %>% 
  rename(planting_site_name=name)
AW_clim_pico_provs<- AW_clim_pico %>% 
  filter(site_type=="provenance") %>% 
  rename(provenance_name = name)

AW_pico_scores_mcmt_shm <- climate_match(AW_clim_pico_provs, 
                                         AW_clim_pico_plantation, 
                                         mcmt_TL=2,
                                         shm_TL = AW_clim_pico_plantation$Normal_1961_1990_SHM/2,
                                         plant_period=c(6190,9120)) %>% 
  select(provenance_name, m_6190, m_9120, everything())
```


##pien
```{r}
AW_clim_pien <- terra::extract(AW_clim_stack, pien_all_vect, method="simple") %>% 
  bind_cols(as.data.frame(pien_all_vect)) %>% 
  bind_cols(crds(pien_all_vect)) 

AW_clim_pien_plantation<- AW_clim_pien %>% 
  filter(site_type=="planting site") %>% 
  rename(planting_site_name=name)
AW_clim_pien_provs<- AW_clim_pien %>% 
  filter(site_type=="provenance") %>% 
  rename(provenance_name = name)

AW_pien_scores_mcmt_shm <- climate_match(AW_clim_pien_provs, 
                                         AW_clim_pien_plantation, 
                                         mcmt_TL=2,
                                         shm_TL = AW_clim_pien_plantation$Normal_1961_1990_SHM/2,
                                         plant_period=c(6190,9120)) %>% 
  select(provenance_name, m_6190, m_9120, everything())
```


##pipo_dixit
```{r}
AW_clim_pipo_dixit <- terra::extract(AW_clim_stack, pipo_dixit_all_vect, method="simple") %>% 
  bind_cols(as.data.frame(pipo_dixit_all_vect)) %>% 
  bind_cols(crds(pipo_dixit_all_vect)) 

AW_clim_pipo_dixit_plantation<- AW_clim_pipo_dixit %>% 
  filter(site_type=="planting site") %>% 
  rename(planting_site_name=name)
AW_clim_pipo_dixit_provs<- AW_clim_pipo_dixit %>% 
  filter(site_type=="provenance") %>% 
  rename(provenance_name = name)

AW_pipo_dixit_scores_mcmt_shm <- climate_match(AW_clim_pipo_dixit_provs, 
                                         AW_clim_pipo_dixit_plantation, 
                                         mcmt_TL=2,
                                         shm_TL = AW_clim_pipo_dixit_plantation$Normal_1961_1990_SHM/2,
                                         plant_period=c(6190,1140)) %>% 
  select(provenance_name, m_6190, m_1140, everything())
```

##pipo_nigro
```{r}
AW_clim_pipo_nigro <- terra::extract(AW_clim_stack, pipo_nigro_all_vect, method="simple") %>% 
  bind_cols(as.data.frame(pipo_nigro_all_vect)) %>% 
  bind_cols(crds(pipo_nigro_all_vect)) 

AW_clim_pipo_nigro_plantation<- AW_clim_pipo_nigro %>% 
  filter(site_type=="planting site") %>% 
  rename(planting_site_name=name)
AW_clim_pipo_nigro_provs<- AW_clim_pipo_nigro %>% 
  filter(site_type=="provenance") %>% 
  rename(provenance_name = name)

AW_pipo_nigro_scores_mcmt_shm_MH <- climate_match(AW_clim_pipo_nigro_provs, 
                                         AW_clim_pipo_nigro_plantation %>% filter(planting_site_name=="MH"), 
                                         mcmt_TL=2,
                                         shm_TL = AW_clim_pipo_nigro_plantation %>% filter(planting_site_name=="MH") %>% pull(Normal_1961_1990_SHM)/2,
                                         plant_period=c(6190,1140)) %>% 
  select(provenance_name, m_6190, m_1140, everything())

AW_pipo_nigro_scores_mcmt_shm_HH <- climate_match(AW_clim_pipo_nigro_provs, 
                                         AW_clim_pipo_nigro_plantation%>% filter(planting_site_name=="HH"), 
                                         mcmt_TL=2,
                                         shm_TL = AW_clim_pipo_nigro_plantation %>% filter(planting_site_name=="HH") %>% pull(Normal_1961_1990_SHM)/2,
                                         plant_period=c(6190,1140)) %>% 
  select(provenance_name, m_6190, m_1140, everything())

```

##pied

```{r}
AW_clim_pied_redmond <- terra::extract(AW_clim_stack, pied_redmond_all_vect, method="simple") %>% 
  bind_cols(as.data.frame(pied_redmond_all_vect)) %>% 
  bind_cols(crds(pied_redmond_all_vect)) 

AW_clim_pied_redmond_plantation<- AW_clim_pied_redmond %>% 
  filter(site_type=="planting site") %>% 
  rename(planting_site_name=name)
AW_clim_pied_redmond_provs<- AW_clim_pied_redmond %>% 
  filter(site_type=="provenance") %>% 
  rename(provenance_name = name)

unique(AW_clim_pied_redmond_plantation$planting_site_name)

AW_pied_redmond_scores_mcmt_shm_NWM <- climate_match(AW_clim_pied_redmond_provs, 
                                         AW_clim_pied_redmond_plantation %>% filter(planting_site_name=="NWM"), 
                                         mcmt_TL=2,
                                         shm_TL = AW_clim_pied_redmond_plantation %>% filter(planting_site_name=="NWM") %>% pull(Normal_1961_1990_SHM)/2,
                                         plant_period=c(6190,1140)) %>% 
  select(provenance_name, m_6190, m_1140, everything())

AW_pied_redmond_scores_mcmt_shm_NWL <- climate_match(AW_clim_pied_redmond_provs, 
                                         AW_clim_pied_redmond_plantation%>% filter(planting_site_name=="NWL"), 
                                         mcmt_TL=2,
                                         shm_TL = AW_clim_pied_redmond_plantation %>% filter(planting_site_name=="NWL") %>% pull(Normal_1961_1990_SHM)/2,
                                         plant_period=c(6190,1140)) %>% 
  select(provenance_name, m_6190, m_1140, everything())

AW_pied_redmond_scores_mcmt_shm_SWH <- climate_match(AW_clim_pied_redmond_provs, 
                                         AW_clim_pied_redmond_plantation%>% filter(planting_site_name=="SWH"), 
                                         mcmt_TL=2,
                                         shm_TL = AW_clim_pied_redmond_plantation %>% filter(planting_site_name=="SWH") %>% pull(Normal_1961_1990_SHM)/2,
                                         plant_period=c(6190,1140)) %>% 
  select(provenance_name, m_6190, m_1140, everything())

AW_pied_redmond_scores_mcmt_shm_SWM <- climate_match(AW_clim_pied_redmond_provs, 
                                         AW_clim_pied_redmond_plantation%>% filter(planting_site_name=="SWM"), 
                                         mcmt_TL=2,
                                         shm_TL = AW_clim_pied_redmond_plantation %>% filter(planting_site_name=="SWM") %>% pull(Normal_1961_1990_SHM)/2,
                                         plant_period=c(6190,1140)) %>% 
  select(provenance_name, m_6190, m_1140, everything())

AW_pied_redmond_scores_mcmt_shm_SWL <- climate_match(AW_clim_pied_redmond_provs, 
                                         AW_clim_pied_redmond_plantation%>% filter(planting_site_name=="SWL"), 
                                         mcmt_TL=2,
                                         shm_TL = AW_clim_pied_redmond_plantation %>% filter(planting_site_name=="SWL") %>% pull(Normal_1961_1990_SHM)/2,
                                         plant_period=c(6190,1140)) %>% 
  select(provenance_name, m_6190, m_1140, everything())

AW_pied_redmond_scores_mcmt_shm_AZM <- climate_match(AW_clim_pied_redmond_provs, 
                                         AW_clim_pied_redmond_plantation%>% filter(planting_site_name=="AZM"), 
                                         mcmt_TL=2,
                                         shm_TL = AW_clim_pied_redmond_plantation %>% filter(planting_site_name=="AZM") %>% pull(Normal_1961_1990_SHM)/2,
                                         plant_period=c(6190,1140)) %>% 
  select(provenance_name, m_6190, m_1140, everything())

AW_pied_redmond_scores_mcmt_shm_AZL <- climate_match(AW_clim_pied_redmond_provs, 
                                         AW_clim_pied_redmond_plantation%>% filter(planting_site_name=="AZL"), 
                                         mcmt_TL=2,
                                         shm_TL = AW_clim_pied_redmond_plantation %>% filter(planting_site_name=="AZL") %>% pull(Normal_1961_1990_SHM)/2,
                                         plant_period=c(6190,1140)) %>% 
  select(provenance_name, m_6190, m_1140, everything())
```

##pimo

```{r}
AW_clim_pimo_urza <- terra::extract(AW_clim_stack, pimo_urza_all_vect, method="simple") %>% 
  bind_cols(as.data.frame(pimo_urza_all_vect)) %>% 
  bind_cols(crds(pimo_urza_all_vect)) 

AW_clim_pimo_urza_plantation<- AW_clim_pimo_urza %>% 
  filter(site_type=="planting site") %>% 
  rename(planting_site_name=name)
AW_clim_pimo_urza_provs<- AW_clim_pimo_urza %>% 
  filter(site_type=="provenance") %>% 
  rename(provenance_name = name)

unique(AW_clim_pimo_urza_plantation$planting_site_name)

AW_pimo_urza_scores_mcmt_shm_PUDH <- climate_match(AW_clim_pimo_urza_provs, 
                                         AW_clim_pimo_urza_plantation %>% filter(planting_site_name=="PUD-H"), 
                                         mcmt_TL=2,
                                         shm_TL = AW_clim_pimo_urza_plantation %>% filter(planting_site_name=="PUD-H") %>% pull(Normal_1961_1990_SHM)/2,
                                         plant_period=c(6190,1140)) %>% 
  select(provenance_name, m_6190, m_1140, everything())

AW_pimo_urza_scores_mcmt_shm_PUDL <- climate_match(AW_clim_pimo_urza_provs, 
                                         AW_clim_pimo_urza_plantation%>% filter(planting_site_name=="PUD-L"), 
                                         mcmt_TL=2,
                                         shm_TL = AW_clim_pimo_urza_plantation %>% filter(planting_site_name=="PUD-L") %>% pull(Normal_1961_1990_SHM)/2,
                                         plant_period=c(6190,1140)) %>% 
  select(provenance_name, m_6190, m_1140, everything())

AW_pimo_urza_scores_mcmt_shm_PUD_PL <- climate_match(AW_clim_pimo_urza_provs, 
                                         AW_clim_pimo_urza_plantation%>% filter(planting_site_name=="PUD_PL"), 
                                         mcmt_TL=2,
                                         shm_TL = AW_clim_pimo_urza_plantation %>% filter(planting_site_name=="PUD_PL") %>% pull(Normal_1961_1990_SHM)/2,
                                         plant_period=c(6190,1140)) %>% 
  select(provenance_name, m_6190, m_1140, everything())
```

#Tool Results

```{r}
##read in results downloaded from tool runs online


```

#Compare match scores
Compare match scores obtained from the 3 methods (ClimateNAr, AdaptWest data download, online webtool)

```{r}


```



